Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
update manuscript

\documentclass{amia}
\usepackage{graphicx}
\usepackage[labelfont=bf]{caption}
\usepackage[superscript,nomove]{cite}
\usepackage{color}
\usepackage{multirow}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

@@

\author{Mehedi Hasan, BS$^{1}$\footnote[1]{Authors provided an equal contribution. \label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Sylvie Naar, PhD$^{2}$, Gwen L. Alexander, PhD$^{3}$, April Idalski Carcone, PhD$^{4}$}

@@

\begin{enumerate}
\item Emails are unstructured text that contains informal information exchange in a non-traditional format.
\item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or collection of sentences. One sentence can be segmented into multiple MI behavior fragments. On the other hand, an MI behavior may comprise several sentences.
\end{enumerate}

@@

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al. applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al. trained Maximum Entropy models for section classification.\cite{tepper2012statistical} Denny et al. proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. \cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries. Liu et al. have shown that a linear chain CRF performed better than the hidden markov model and Maxent on the NIST sentence boundary detection.\cite{liu2005using} Recently, Treviso et al. proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word vectors,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.

@@

@@

The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We formulate the segmentation task as a binary classification problem where each word or punctuation mark annotated with a label 1 or 0, to indicate whether it precedes a new segment or not. In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning method, CRF and deep learning methods, such as MLP, BRNN and CRNN. For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current wordor punctuation. Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word or punctuation precedes a new segment or not. For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary. For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words. Trained word vectors were also utilized when words or punctuation marks are not available in the pre-trained word2vec word vectors. Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models. It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together. Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study. CRF utilized lexical features which represent each word in an e-Coaching email.

@@

\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is a discriminative model of the conditional distribution of the entire sequence of labels given the observation sequence. A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq. \ref{Eq:crf}

@@

Here $Z_x$ is a normalization factor over all possible labelings of $x$, and $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$. The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm. In our experiments, the following features were utilized with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) whether the word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is a title word or not? (e.g. ``The'' is a title word but ``the'' is not) v) POS tags

\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase. \cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models. For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors. Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$. Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.

@@

E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text. A pooling operation was performed over time to find the most significant features in a region. After applying $n_f=100$ different filters on convolutional layer, $m \times n_f$ data are fed into bidirectional recurrent neural networks, which contain 200 hidden units. Then, dropout is used to randomly drop 50\% of the connections and passes them through a fully connected layer, where a sigmoid operation is performed, provide the probability of whether the word or punctuation precedes a new segment or not. We experimentally determined that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting. To utilize our source codes by other research studies, source codes of all experiments will be publicly available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

@@

\section*{Results}
Our experimental results have two important directions. First, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}. Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.\\

\begin{table}[ht]
\centering
\caption{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used. The highest value for each performance metric
is highlighted in bold.}
\label{tab:result_base}
  \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.782 & 0.691 & 0.733 & 0.983 & 0.984 & 0.984 \\ \hline
 MLP & \textbf{0.836} & 0.593 & 0.694 & 0.982 & 0.983 & 0.982 \\ \hline
 BRNN & 0.606 & 0.680 & 0.641 & 0.977 & 0.976 & 0.976 \\ \hline
 CRNN & 0.775 & \textbf{0.797} & \textbf{0.785} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986} \\ \hline
  \end{tabular}
\end{table}

As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection. CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes. BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure. On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''. CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes. Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified. For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes.\\

\begin{table}[ht]
\centering
\caption{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together. The highest value for each performance metric
is highlighted in bold.}
\label{tab:result_weighted_avg}
 \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
 MLP & \textbf{0.814} & 0.732 & 0.771 & 0.987 & 0.987 & 0.987 \\ \hline
 BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
 CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
  \end{tabular}
\end{table}

\section*{Discussion}
This study is the first effort to evaluate the automatic segmentation of e-Coaching emails. Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study. CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''. The robust performance of CRNN provides the evidence that deep learning models are capable to learn MI behaviors from the clinical exchange. It also indicates that punctuation and POS features are important with word embeddings or lexical features for all machine learning methods are employed. Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.

Punctuation mark and POS features made a significant improvement in the performance of machine learning and all deep learning methods. In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features. This results also indicate that segmentation performance might be improved by adding additional relevant features.

The convolutional layer made a significant difference between the performance of CRNN and BRNN on email segmentation. CRNN demonstrates 22.5\% and 10.7\% higher F1-Measure in ``new segment'' detection and 1\% and 0.5\% higher F1-Measure in overall compared to BRNN when word embeddings and all features are used, respectively. \\

@@

\begin{table}[ht]
\centering
\caption{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used. Highest value for each experiment across all models is highlighted in boldface.}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
 & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings or lexical features & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}

>>>>

\documentclass{amia}
\usepackage{graphicx}
\usepackage[labelfont=bf]{caption}
\usepackage[superscript,nomove]{cite}
\usepackage{color}
\usepackage{multirow}
\renewcommand*{\thefootnote}{\alph{footnote}}

@@

\author{Mehedi Hasan, BS$^{1}$\footnote{Authors provided an equal contribution. \label{footnote1}}, Alexander Kotov, PhD$^{1}$\textsuperscript{\ref{footnote1}}, Sylvie Naar, PhD$^{2}$, Gwen L. Alexander, PhD$^{3}$, April Idalski Carcone, PhD$^{4}$}

@@

\begin{enumerate}
\item Emails are unstructured text that contains informal information exchange in a non-traditional format. However, a traditional MI transcript corresponds to an MI interview
session, contains a dialogue between a counselor and a patient.
\item Discourse segments in e-Coaching emails do not necessarily correspond to sentences or collection of sentences. One sentence can be segmented into multiple MI behavior fragments. On the other hand, an MI behavior may comprise several sentences.
\end{enumerate}

@@

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers of electronic health records (EHR)\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection of EHR.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al. applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al. trained Maximum Entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al. proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. \cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries in EHR. Liu et al. have shown that a linear chain CRF performed better than the hidden markov model and Maxent in sentence boundary detection, evaluated using the official
NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al. proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.

@@

Our proposed approach is novel for the segmentation of e-Coaching emails since previous studies mainly focused on the segmentation of EHR documents into sections, headers and sentences. However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually. Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of clinical exchanges.

@@

The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We formulate the segmentation task as a binary classification problem where each word or punctuation mark annotated with a label 1 or 0, to indicate whether it precedes a new segment or not. In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation. Each sample is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not. For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} In our experiment, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{https://code.google.com/p/word2vec/} When words or punctuation marks are not available in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus referred to trained word embeddings. Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models. It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together. Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study. CRF utilized lexical features which represent each word in an e-Coaching email.

@@

\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is modeled with an exponential distribution provides the conditional distribution of the entire output sequence given the observation sequence. A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq. \ref{Eq:crf}

@@

Here $Z_x$ is a normalization factor over all possible labelings of $x$, and $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$. The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm. In our experiments, the following features were utilized with CRF models: i) current word or punctuation ii) next and previous 3 words or punctuations iii) binary feature indicating the word or punctuation is a special character (';', '?', '.', ',', '!', ':', etc) or not iv) binary feature indicating the word is a title word or not? (e.g. ``The'' is a title word but ``the'' is not) v) POS tags

\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multi layer perceptron with a single hidden layer. In an MLP networks, there are no cycles or loops where the information moves forward only, from the input nodes through the hidden nodes and to the output nodes. This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase. \cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/mlp.jpeg}
    \caption{\textbf{A multi layer perceptron with a single hidden layer}}
    \label{fig:mlp}
\end{figure}

@@

E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The primary purpose of convolution is to extract new features depending on neighbor words. We used 1D convolution which simply moves one step at a time on vertically. In this layer, one filter is responsible for extraction of one feature. After applying $n_f$ different filters with zero-padding on both sides of the text, $n_f$ useful features are produced in the convolution layer for each word. A max pooling operation was then performed over time to find the most significant features in a text region. The bidirectional recurrent layer receives new features extracted in the convolution layer. RNN is usually used to capture the previous history. Moreover, bidirectional RNN is capable of capturing both previous and past histories because bidirectional RNN utilized the features by looking at forward and backward states. The purpose of the fully connected layer is to use the bidirectional RNN layer output for classifying each word or punctuation into ``new segment'' and ``same segment'' classes. Since a fully connected layer has a greater number of parameters, they are more likely excessively co-adapt these parameters causing over-fitting. To prevent this over-fitting problem, dropout is used to randomly drop 50\% of the connections on the fully connected layer. After that, a sigmoid operation is performed on the output layer to provide the probability of whether the word or punctuation precedes a new segment or not. We experimentally determined the optimal parameters and found that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, number of filters is 100, max-pool size is 3, convolutional layer activation is $relu$, RNN layer activation is $tanh$ and number of recurrent units in RNN is 200. Adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is applied with 50 epochs when the batch size is 32 and the learning rate is 0.001. To utilize our source codes by other research studies, source codes of all experiments will be publicly available\footnote{https://github.com/teanalab/xxxxx} upon acceptance of this paper for publication.

@@

\section*{Results}
Our experimental results have two important directions. First, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}. Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.

For a given sample with $2l$ words or punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors. Figure~\ref{fig:length-mlp} demonstrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$. Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.

As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection. CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes. BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure. On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''. CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes. Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified. For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection. \\

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used. The highest value for each performance metric is highlighted in bold.}}
\label{tab:result_base}
  \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.782 & 0.691 & 0.733 & 0.983 & 0.984 & 0.984 \\ \hline
 MLP & \textbf{0.836} & 0.593 & 0.694 & 0.982 & 0.983 & 0.982 \\ \hline
 BRNN & 0.606 & 0.680 & 0.641 & 0.977 & 0.976 & 0.976 \\ \hline
 CRNN & 0.775 & \textbf{0.797} & \textbf{0.785} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986} \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_roc} shows the impact of individual fetures as well as their combination on the segmentation task. Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.\\

\begin{table}[ht]
\centering
\caption{\textbf{AUC values of all classifiers when only word embeddings or lexical features and combination of word embeddings or lexical, punctuation and POS features are used. Highest value for each experiment across all models is highlighted in boldface.}}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{AUC}} \\\cline{2-5}
 & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 word embeddings + POS & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 word embeddings + punctuation & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN methods for identification of ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes when all features are used together. The highest value for each performance metric
is highlighted in bold.}}
\label{tab:result_weighted_avg}
 \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
 MLP & \textbf{0.814} & 0.732 & 0.771 & 0.987 & 0.987 & 0.987 \\ \hline
 BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
 CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
  \end{tabular}
\end{table}

\section*{Discussion}
This study is the first effort to evaluate the automatic segmentation of e-Coaching emails. Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study. CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''. The robust performance of CRNN provides the evidence that deep learning models are capable to learn the hidden characteristics of MI behaviors from the clinical exchange. It also indicates that punctuation and POS features are important with word embeddings or lexical features for all machine learning methods are employed. Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.

Punctuation mark and POS features made a significant improvement in the performance of machine learning and all deep learning methods. Specially, punctuation features have the higher individual impact on model performance compared to POS features. In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features. This results also indicate that segmentation performance might be improved by adding additional relevant features.

The convolutional layer made a significant difference between the performance of CRNN and BRNN on email segmentation. CRNN demonstrates 22.5\% and 10.7\% higher F1-Measure in ``new segment'' detection and 1\% and 0.5\% higher F1-Measure in overall compared to BRNN when word embeddings and all features are used, respectively. In CRNN, convolution layer performed a series of convolutions and pooling operations during which a number of high-level important features are extracted from the word embeddings of e-Coaching emails. These high-level features are then utilized by the bidirectional RNN of CRNN model, which provide an improved performance on email segmentation. On the other hand, traditional BRNN model received words as input features and their word embeddings are directly utilized in the input layer.

@@

In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance. In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets. AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets,\cite{hu2015kernelized} which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}.
